---
title: "case_EDA"
author: "MDS"
date: "March 29, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####################
 Explore Data Frame
###################

```{r library, echo=F}
library(ggplot2)
library(dplyr)
```

Exploring Categorical Data 

Reading Data From .CSV:

```{r Read Data, echo=FALSE}
comics <- read.csv('comics.csv')
comics[1,]
```

1.Print the first 5 rows of the data using head()
2.Check the structure of the data

```{r Read Data(2), echo=F}
head(comics, 5)
str(comics)
```

Check levels

```{r levels, echo=F}
# Check levels of align
levels(comics$align)
# levels(comics$gender)
```

Create a 2-way contingency table: tab

```{r table, echo=F}
tab <- table(comics$align,comics$gender )
tab
```


Bar Chart

```{r bar_chart,echo=F}
ggplot(comics, aes(x = gender, fill = align)) +
  geom_bar()
```

Filtering

```{r filter, echo=f}
# Filter the data that has the fewest total of 'Align' entries
comics_filtered <- comics %>%
  filter(align != 'Reformed Criminals') %>%
  droplevels()

levels(comics_filtered$align)
```

Create side-by-side barchart of alignment by gender

```{r Side_bar_chart, echo=F}
ggplot(comics, aes(x = align, fill = gender)) + 
  geom_bar(position = 'dodge') +
  theme(axis.text.x = element_text(angle = 90))

ggplot(comics, aes(x = gender, fill = align)) + 
  geom_bar(position = 'dodge') +
  theme(axis.text.x = element_text(angle = 90))
```

From counts to proportions

```{r proportion}
prop.table(tab,1) # condition on row
prop.table(tab,2) # condition on column
```

Stack Bar Chart

```{r stack_bar_chart, echo=FALSE}
ggplot(comics, aes(x = gender, fill = align)) +
  geom_bar(position = "fill")

ggplot(comics, aes(x = align, fill = gender)) +
  geom_bar(position = "fill")
```

Histogram, to check the distribution

```{r histogram, echo=F}
ggplot(comics, aes(x = align)) + 
  geom_text(stat = "count",
            aes(label = ..count..), vjust = -1) +
  geom_bar()
```

Faceting: Plot of alignment broken down by gender

```{r facet, echo=F}
ggplot(comics, aes(x = align)) + 
  geom_bar() +
  facet_wrap(~ gender)
```

Exploring Numerical Data

```{r cars, echo=F}
cars <- read.csv('cars04.csv')
head(cars)
str(cars)
```

Plot a histogram of city_mpg faceted by suv, a logical variable indicating whether the car is an SUV or not. 

```{r plot_car, echo=F}
ggplot(cars, aes(x = city_mpg)) +
  geom_histogram() + 
  facet_wrap(~ suv)
```

A quick look at unique(cars$ncyl) shows that there are more possible levels of ncyl than you might think. Here, restrict your attention to the most common levels.

```{r check,echo=F}
#histogram
ggplot(cars, aes(x = ncyl)) +
  geom_histogram()

#filter
common_cyl <- filter(cars, ncyl %in% c(4, 6, 8))

#boxplot
ggplot(common_cyl, aes(x = as.factor(ncyl), y = city_mpg)) +
  geom_boxplot()

#density
ggplot(common_cyl, aes(x = city_mpg, fill = as.factor(ncyl))) +
  geom_density(alpha = .3)
```

Create a histogram of the distribution of horsepwr across all cars and add an appropriate title. Start by piping in the raw dataset.

```{r histogram, eccho=F}
cars %>%
  ggplot(aes(horsepwr)) +
  geom_histogram() +
  ggtitle('horsepwr')
```

Create a second histogram of the distribution of horsepower, but only for those cars that have an msrp less than $25,000.Keep the limits of the x-axis so that they're similar to that of the first plot. Experiment with the bin and add a descriptive title.

```{r play_bin, eccho=F}
cars %>% 
  filter(msrp<25000) %>%
  ggplot(aes(horsepwr)) +
  geom_histogram(bins=50) +
  xlim(c(90, 550)) +
  ggtitle('hrspwr')
```

```{r}
# histogram
cars %>%
  ggplot(aes(msrp)) +
  geom_histogram()

# Construct box plot of msrp
cars %>%
  ggplot(aes(x = 1, y = msrp)) +
  geom_boxplot()

# Exclude outliers from data
cars_no_out <- cars %>%
  filter(msrp<100000)

# Construct box plot of msrp using the reduced dataset
cars_no_out %>%
  ggplot(aes(x = 1, y = msrp)) +
  geom_boxplot()
```

Summarize Numerical Data

```{r}
gapminder <- read.csv('gapminder.csv')

# Create a dataset called gap2007 that contains only data from the year 2007.
gap2007 <- filter(gapminder, year==2007)
```

group_by and summarize

```{r}
# Using gap2007, calculate the mean and median life expectancy for 
# each country. Don't worry about naming the new columns produced by summarize().
gap2007 %>%
  group_by(cont) %>%
  summarize(mean(life_exp),
            median(life_exp),
            var(life_exp),
            sd(life_exp))

# Confirm the trends that you see in the medians by 
# generating side-by-side box plots of life expectancy for each continent.
gap2007 %>%
  ggplot(aes(x = cont, y = life_exp, fill=cont)) +
  geom_boxplot()
```

```{r}
# Checking distribution
gap2007 %>%
  ggplot(aes(x = life_exp, fill=cont)) +
  geom_density(alpha=.3)

gap2007 %>% 
  ggplot(aes(x = cont, y = life_exp, fill=cont)) +
  geom_boxplot()

# Filtering data
# Filter for Asia, add column indicating outliers
gap_asia <- gap2007 %>%
  filter(cont == "Asia") %>%
  mutate(is_outlier = life_exp < 50)

# Remove outliers, create box plot of lifeExp
gap_asia %>%
  ggplot(aes(x = life_exp, fill=cont)) +
  geom_density(alpha=.3)

gap_asia %>% 
  filter(!is_outlier) %>%
  ggplot(aes(x = 1, y = life_exp)) +
  geom_boxplot()
```

